---
title: "Generate test set for XGB models"
author: "Kaspar Bresser"
date: "12/01/2021"

output: 
  github_document:
    toc: true
  html_document: 
    theme: simplex
    highlight: pygments
#    code_folding: show
    self_contained: TRUE
    toc: yes
    toc_float:
      collapsed: no
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      autodep  = TRUE,
                      cache = FALSE,
#                     comment = NA,
                      fig.width = 5,
                      fig.asp = 0.618,
                      fig.align = "center")
```

```{r startup}
library(here)
library(tidyverse)
```




First, import the ligands that we'll use as positive cases. 

```{r read_ligands}
(ligand.table <- read_tsv(here("Output", "Broad_peptides_subsample.tsv")))
```

Import the swissprot IDs that are in the feature library

```{r import_IDs}
here( "Protein_per_Uniprot_entry_library_v2_RBP_GC_length_codon_AA_m6A_m5C_AtoI_m1A_m7G_CD8miRDB_PTM.csv") %>% 
  read_tsv() %>% 
  pull(Entry) -> IDs.in.library
```



Next, define a function that will take the netMHCpan output table for a given allele, plus the allele name. This function will read the file, add the allele name as a column, annotate which peptides are positive cases, and finally subset the dataset to 350 positive and 350,000 negative cases. 

```{r define_import_function}

import_peptides <- function(file, al){

  
  ligand.table %>% 
    filter(allele == al) %>% 
    pull(sequence) -> detected.ligands
  
  file %>% 
    read_tsv(skip = 1) %>% 
    filter(ID %in% IDs.in.library) %>% 
    mutate(allele = al) %>% 
    select( c("Peptide", "ID", "Rank", "allele")) %>% 
    mutate(ligand = case_when(Peptide %in% detected.ligands ~ TRUE,
                              TRUE ~ FALSE)) %>% 
    group_by(ligand) %>% 
    nest() %>% 
    ungroup() %>% 
    mutate(data = map2(data, c(350000, 350), sample_n)) %>% 
    unnest(data)
      
}
```

Now, get the the files, set the alleles as names, apply the `import_peptides` function and row bind the output into a single table. 

```{r import_peptides}
file.names <- list.files(here("Broad_Data", "HLA_peptidomics", "netMHCpan"), pattern = "broad")

file.names %>% 
  here("Broad_Data", "HLA_peptidomics", "netMHCpan", .) %>% 
  set_names(str_remove(file.names, "_broad.txt")) -> files

files

files %>% 
  map2(., names(.), import_peptides) %>% 
  purrr::reduce(bind_rows) -> test.table

test.table
```

Now we'll add the netChop processing scores. I'll take an old file for which I applied the neolution pipeline (which spits out the processing scores) on all swissprot annotated proteins. 

```{r add_processing_scores}
here("Broad_Data", "HLA_peptidomics", "swissprot_A0201_9mer.csv") %>% 
  read_csv() %>% 
  transmute( Peptide = peptide, processing_score = processing_score) %>%
  distinct(Peptide, .keep_all = T) %>% 
  inner_join(test.table, by = "Peptide") -> test.table

test.table
```

Finally add the RNAseq data. Make sure to z-score normalize

```{r add_rna}
(TPM.table <- read_tsv(here("Broad_Data", "RNAseq", "averaged_TPM_values_salmon.tsv")))

TPM.table %>% 
  filter(TPM > 0) %>% 
  transmute(ID = swissprot_id, rna = TPM) %>% 
  distinct(ID, .keep_all = T) %>% 
  mutate(rna = scale(log10(rna))[,1]) %>% 
  inner_join(test.table, by = "ID") -> test.table

test.table

test.table %>% 
  ggplot(aes(x = allele, y = rna, fill = ligand))+
  geom_boxplot()
```

And write out the file

```{r write_out}
test.table %>% 
  write_tsv(here("Output", "Broad_test_set_forXGB1.tsv"))


```

