---
title: "XGB results top 0.1% Broad data"
author: "Kaspar Bresser"
date: "11/01/2021"

output: 
  github_document:
    toc: true
  html_document: 
    theme: simplex
    highlight: pygments
#    code_folding: show
    self_contained: TRUE
    toc: yes
    toc_float:
      collapsed: no
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      autodep  = TRUE,
                      cache = FALSE,
#                     comment = NA,
                      fig.width = 5,
                      fig.asp = 0.618,
                      fig.align = "center")
```

Used the analysis below to compare the performance of the XGB models.

```{r loading}
library(here)
library(bayestestR)
library(tidyverse)
```

## Import and tidy data

Import the results table.

```{r import_results}
here("Output", "XGB") %>% 
  list.files() %>% 
  here("Output", "XGB", .) %>% 
  map(read_tsv) %>% 
  reduce(bind_rows)  -> XGB.results

XGB.results
```

Convert the ligand column to a binary so it can be used for cumsum calculation, pivot to longer data by putting the models in a grouped format.

```{r prep_data}
XGB.results %>% 
  mutate(detected = case_when(ligand == TRUE ~ 1,
                              TRUE ~ 0)) %>% 
  pivot_longer(cols = !c("detected", "ligand", "allele"), 
               names_to =  "model", 
               values_to =  "score") -> XGB.results

XGB.results
```


For these analysis we'll focus on the top 0.1% scoring peptides for each model. Let's subset on those. Note that I arrange on both both model scores and a random number for tie breaking.

```{r data_prep3}
XGB.results %>% 
  group_by(allele, model) %>% 
  mutate(random = sample(1:n())) %>% 
  arrange(desc(score), random, .by_group = T) %>% 
  slice_head(prop = .001) -> XGB.results
 
XGB.results
```


Next we'll calculate the cumulative sums by grouping by allele and model. 

Note that `cumsum()` takes an ordered dataframe. 

```{r prep_data2}
XGB.results %>% 
  group_by(allele, model) %>%
  mutate(peptides = 1:n(), 
         detected_ligands = cumsum(detected)) %>% 
  ungroup() -> XGB.results

XGB.results
```


Lastly, set the ordering in which we'd like the models to appear in plots, by converting the models variable to a factor

```{r data_prep4}
unique(XGB.results$model)

#model.order <- c("aff_only", "aff_chop_only",  "aff_chop_Library","aff_chop_rna",  "aff_chop_Library_rna")

model.order <- c("aff_only", "aff_chop_only",  "Library_chop_aff","rna_aff_chop",  "Library_rna_chop_aff")

allele.order <- c("A2402", "A0201", "B3501", "B5101", "A1101", "A3101", "B4001", "B0702")

colors <- c("#27aae1", "#6675b8", "#834e9f", "#2bb673", "#2e7a3c")

XGB.results %>% 
  filter(model %in% model.order) %>% 
  mutate(model = as_factor(model),
         model = fct_relevel(model, model.order),
         allele = as_factor(allele),
         allele = fct_relevel(allele, allele.order)) -> XGB.results

XGB.results
```


## Positive predictive value

Lets first plot the positive predictive value (PPV), i.e. the number of true positives within a certain threshold. Calculate the PPV as the number of true positives divided by the total number of peptides within the selected pool.

```{r PVV_plot_increment, fig.width=10}
XGB.results %>%  
  group_by(allele, model) %>% 
  summarise(PPV = sum(detected)/n()) %>% 
   ggplot( aes(x = allele, y = PPV, fill = model)) +
    geom_bar(stat = "identity", width=0.75, position=position_dodge(width=0.85), color = "black")+
    scale_fill_manual(values = colors)+
    ggtitle("PPV in top 0.1%")+
    theme_classic()+
    theme(plot.title = element_text(hjust = 0.5), panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_Ext_PPV_bar_final.pdf"), width = 4, height = 2, scale = 2)
```

Can also plot the PPV on a continuous scale across the top 0.1%. 

```{r plot_peptides_found_continuous, fig.width=8, fig.asp=.9}
XGB.results %>% 
  filter(model %in% model.order) %>% 
    ggplot(aes(x = peptides, y = detected_ligands/peptides, color = model))+
    geom_line(size = 1)+
    scale_color_manual(values = colors)+
    facet_wrap(~allele, scales = "free", nrow = 4)+
    theme_classic()+
    theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_Ext_PPV_line_final.pdf"), width = 4, height = 4, scale = 2)
```

## Cumulative sums

We can additionally plot the cumulative sums of all models.

```{r plot_peptides_found, fig.width=8, fig.asp=.9}
XGB.results %>% 
  filter(model %in% model.order) %>% 
  ggplot(aes(x = peptides, y = detected_ligands, color = model))+
  geom_line(size = 1)+
  scale_color_manual(values = colors)+
  facet_wrap(~allele, scales = "free", nrow = 4)+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_Ext_cumsum_final.pdf"), width = 4, height = 4, scale = 2)
```

## Area under the curve

Or calculate the AUC of these curves as a summary metric and plot that. Used the `area_under_curve()` function from the [bayestestR](https://easystats.github.io/bayestestR/index.html) package. 

```{r plot_auc, fig.width=10, fig.asp=.45}
XGB.results %>% 
  group_by(allele, model) %>% 
  summarise(auc = area_under_curve(peptides, peptides)) %>% 
  distinct(auc)

XGB.results %>% 
  group_by(allele, model) %>% 
  summarise(auc = area_under_curve(peptides, detected_ligands) / 61249.5) %>% 
  ggplot( aes(x = allele, y = auc, fill = model)) +
  geom_bar(stat = "identity", width=0.75, position=position_dodge(width=0.85), color = "black")+
  scale_fill_manual(values = colors)+
  ggtitle("AUC in top 0.1%")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_Ext_AUC_final.pdf"), width = 5, height = 2, scale = 2)
```
