---
title: "Test XGB models"
author: "Kaspar Bresser"
date: "24/12/2021"

output: 
  github_document:
    toc: true
  html_document: 
    theme: simplex
    highlight: pygments
#    code_folding: show
    self_contained: TRUE
    toc: yes
    toc_float:
      collapsed: no
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      autodep  = TRUE,
                      cache = FALSE,
#                     comment = NA,
                      fig.width = 5,
                      fig.asp = 0.618,
                      fig.align = "center")
```



```{r loading}
library(here)
library(caret)
library(xgboost)
library(tidyverse)
```

Import the feature library

```{r add_features}
feature.table <- read_tsv(here("Data", "Protein_per_Uniprot_entry_library_v2_RBP_GC_length_codon_AA_m6A_m5C_AtoI_m1A_m7G_CD8miRDB_PTM.csv")) %>% 
  mutate(across(everything(), replace_na, 0))

feature.table <- rename(feature.table, swissprot_id = Entry)
```


Get the file list, extract the names that will be used as column names


```{r get_files}
files <- list.files(here("XGB_models_final"), pattern = "xgb")
#files <- c("xgb_ligands|aff_chop_only|11_01_2022.RDS", "xgb_ligands|rna_aff_chop|11_01_2022.RDS")



files %>% 
  str_extract("\\|.*\\|") %>% 
  str_remove_all("\\|") -> names

files <- set_names(files, names)

files
```

Import the XGB models.

```{r get_models}
files %>% 
  map(~read_rds(here("XGB_models_final", .))) %>% 
  set_names(names) -> XGB.models

XGB.models


```

get column names.

```{r read_column_names}
here("Output", "test_train_tables_new", "rf_test_peptides.tsv") %>% 
  read_tsv( n_max = 1) %>% 
  names() -> column.names
```


```{r read_rna}
here("Data", "rna_seq_data", "all_lines_final.rds") %>% 
  read_rds() %>% 
  unnest(expr) -> expr.data
```


Will import the test set in chunks, perform the predictions with each model and write out the scores in a single pipe to keep from depleting memory. The test set contains 2,540,538 entries. 

```{r get_starts}
2540538/33

2540538-76986

seq(0, 2463552, 76986)

start.points <- seq(0, 2463552, 76986)+1
```

Define a function that will import a portion of the test data and join with the feature library.

Next, loop over the start points. In each loop use the subset of the test data to perform predictions with all imported models. Column-bind this to the ligand column of the same partition of the data, and write out.

```{r predictions}
importstuff <- function(start){
  here("Output", "test_train_tables_new", "rf_test_peptides.tsv") %>% 
    read_tsv(skip = start , n_max = 76986, col_names = column.names) %>% 
    mutate(tumor = fct_recode(tumor, M026.X1 = "MEL1", SKMEL95 = "MEL2", NKIRTIL006 = "MEL3")) %>% 
    dplyr::select(-rna) %>% 
    left_join(expr.data, by = c("tumor", "swissprot_id")) %>% 
    inner_join(feature.table)  -> out
  return(out)
}


for (i in start.points){
    XGB.models %>% 
      map2_dfc(list(importstuff(i)) , ~predict(object = .x, newdata = .y, type = "prob")$`TRUE`) %>% 
      bind_cols(read_tsv(here("Output", "test_train_tables_new", "rf_test_peptides.tsv"),
                         skip = i, 
                         n_max = 76986, 
                         col_names = column.names, 
                         col_select = c(tumor, ligand))) %>% 
      write_tsv(here("Output", "XGB_final", paste0("prediction_results_", i, ".tsv")))
  gc()
}
```





