---
title: "XGB results top 0.1%"
author: "Kaspar Bresser"
date: "28/12/2021"

output: 
  github_document:
    toc: true
  html_document: 
    theme: simplex
    highlight: pygments
#    code_folding: show
    self_contained: TRUE
    toc: yes
    toc_float:
      collapsed: no
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      autodep  = TRUE,
                      cache = FALSE,
#                     comment = NA,
                      fig.width = 5,
                      fig.asp = 0.618,
                      fig.align = "center")
```

Used the analysis below to compare the performance of the XGB models.

```{r loading}
library(here)
library(bayestestR)
library(tidyverse)
```

## Import and tidy data

Import the results table.

```{r import_results}
here("Output", "XGB_final") %>% 
  list.files() %>% 
  here("Output", "XGB_final", .) %>% 
  map(read_tsv) %>% 
  reduce(bind_rows) %>% 
  mutate(tumor = fct_recode(tumor, M026.X1 = "M026", SKMEL95 = "mel95", NKIRTIL006 = "RTE")) -> XGB.results

XGB.results

XGB.results %>% 
  count(ligand, tumor)
```

Convert the ligand column to a binary so it can be used for cumsum calculation, pivot to longer data by putting the models in a grouped format.

```{r prep_data}
XGB.results %>% 
  mutate(detected = case_when(ligand == TRUE ~ 1,
                              TRUE ~ 0)) %>% 
  pivot_longer(cols = !c("detected", "ligand", "tumor"), 
               names_to =  "model", 
               values_to =  "score") -> XGB.results

XGB.results
```


For these analysis we'll focus on the top 0.1% scoring peptides for each model. Let's subset on those. Note that I arrange on both both model scores and a random number for tie breaking.

```{r data_prep3}
XGB.results %>% 
  group_by(tumor, model) %>% 
  mutate(random = sample(1:n())) %>% 
  arrange(desc(score), random, .by_group = T) %>% 
  slice_head(prop = .001) -> XGB.results
 
XGB.results
```


Next we'll calculate the cumulative sums by grouping by tumor and model. 

Note that `cumsum()` takes an ordered datafame. 

```{r prep_data2}
XGB.results %>% 
  group_by(tumor, model) %>%
  mutate(peptides = 1:n(), 
         detected_ligands = cumsum(detected)) %>% 
  ungroup() -> XGB.results

XGB.results
```


Lastly, set the ordering in which we'd like the models to appear in plots, by converting the models variable to a factor

```{r data_prep4}
unique(XGB.results$model)

model.order <- c("aff_only", "aff_chop_only", "aff_chop_Library", "aff_chop_rna", 
                 "aff_chop_ribo", "aff_chop_rna_ribo", "aff_chop_Library_rna", "aff_chop_Library_ribo",
                 "aff_chop_Library_rna_ribo")

XGB.results %>% 
  mutate(model = as_factor(model),
         model = fct_relevel(model, model.order)) -> XGB.results

XGB.results
```


## Positive predictive value

Lets first plot the positive predictive value (PPV), i.e. the number of true positives within a certain threshold. Calculate the PPV as the number of true positives divided by the total number of peptides within the selected pool.

```{r PVV_plot_bars_all, fig.width=6}
XGB.results %>%  
  group_by(tumor, model) %>% 
  summarise(PPV = sum(detected)/n()) %>% 
ggplot( aes(x = tumor, y = PPV, fill = model)) +
  geom_bar(stat = "identity", width=0.75, position=position_dodge(width=0.85), color = "black")+
  ggtitle("PPV in top 0.1%")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_PPV_bar_all_final.pdf"), width = 4, height = 2, scale = 2)
```

Plot improvement gained with library:

```{r PVV_plot_bars_core, fig.width=6}
core.models <- c("aff_only", "aff_chop_only", "aff_chop_Library")

XGB.results %>%  
  filter(model %in% core.models) %>% 
  group_by(tumor, model) %>% 
  summarise(PPV = sum(detected)/n()) %>% 
ggplot( aes(x = tumor, y = PPV, fill = model)) +
  geom_bar(stat = "identity", width=0.75, position=position_dodge(width=0.85), color = "black")+
  ggtitle("PPV in top 0.1%")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_PPV_bar_core_final.pdf"), width = 3, height = 2, scale = 2)
```

Compare with expression

```{r PVV_plot_bars_expression, fig.width=6}
expr.models <- c("aff_chop_only", "aff_chop_Library","aff_chop_rna", 
                 "aff_chop_ribo", "aff_chop_rna_ribo" )

XGB.results %>%  
  filter(model %in% expr.models) %>% 
  group_by(tumor, model) %>% 
  summarise(PPV = sum(detected)/n()) %>% 
ggplot( aes(x = tumor, y = PPV, fill = model)) +
  geom_bar(stat = "identity", width=0.75, position=position_dodge(width=0.85), color = "black")+
  ggtitle("PPV in top 0.1%")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_PPV_bar_expression_final.pdf"), width = 3, height = 2, scale = 2)
```

And the combined:

```{r PVV_plot_bars_combined, fig.width=6}
comb.models <- c("aff_chop_only", "aff_chop_Library","aff_chop_Library_rna", "aff_chop_Library_ribo",
                 "aff_chop_Library_rna_ribo" )

XGB.results %>%  
  filter(model %in% comb.models) %>% 
  group_by(tumor, model) %>% 
  summarise(PPV = sum(detected)/n()) %>% 
ggplot( aes(x = tumor, y = PPV, fill = model)) +
  geom_bar(stat = "identity", width=0.75, position=position_dodge(width=0.85), color = "black")+
  ggtitle("PPV in top 0.1%")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_PPV_bar_combined_final.pdf"), width = 3, height = 2, scale = 2)
```




Can also plot the PPV on a continuous scale across the top 0.1%. 

```{r PPV_continuous_all, fig.width=12, fig.asp=.4}
XGB.results %>% 
  filter(model %in% model.order) %>% 
  ggplot(aes(x = peptides, y = detected_ligands/peptides, color = model))+
  geom_line(size = 1)+
  facet_wrap(~tumor, scales = "free")+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_PPV_line_all_final.pdf"), width = 6, height = 2, scale = 2)
```

Core models:

```{r PPV_continuous_core, fig.width=12, fig.asp=.4}
XGB.results %>% 
  filter(model %in% core.models) %>% 
  ggplot(aes(x = peptides, y = detected_ligands/peptides, color = model))+
  geom_line(size = 1)+
  facet_wrap(~tumor, scales = "free")+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_PPV_line_core_final.pdf"), width = 6, height = 2, scale = 2)
```


```{r PPV_continuous_expr, fig.width=12, fig.asp=.4}
XGB.results %>% 
  filter(model %in% expr.models) %>% 
  ggplot(aes(x = peptides, y = detected_ligands/peptides, color = model))+
  geom_line(size = 1)+
  facet_wrap(~tumor, scales = "free")+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_PPV_line_expr_final.pdf"), width = 6, height = 2, scale = 2)
```


```{r PPV_continuous_comb, fig.width=12, fig.asp=.4}
XGB.results %>% 
  filter(model %in% comb.models) %>% 
  ggplot(aes(x = peptides, y = detected_ligands/peptides, color = model))+
  geom_line(size = 1)+
  facet_wrap(~tumor, scales = "free")+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_PPV_line_expr_final.pdf"), width = 6, height = 2, scale = 2)
```

## Cumulative sums

We can additionally plot the cumulative sums of all models.

```{r plot_cumsum_all, fig.width=12, fig.asp=.4}
XGB.results %>% 
  filter(model %in% model.order) %>% 
  ggplot(aes(x = peptides, y = detected_ligands, color = model))+
  geom_line(size = 1)+
  facet_wrap(~tumor, scales = "free")+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_cumsum_all_final.pdf"), width = 6, height = 2, scale = 2)
```


Core models:

```{r plot_cumsum_core, fig.width=12, fig.asp=.4}
XGB.results %>% 
  filter(model %in% core.models) %>% 
  ggplot(aes(x = peptides, y = detected_ligands, color = model))+
  geom_line(size = 1)+
  facet_wrap(~tumor, scales = "free")+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_cumsum_core_final.pdf"), width = 6, height = 2, scale = 2)
```

```{r plot_cumsum_expr, fig.width=12, fig.asp=.4}
XGB.results %>% 
  filter(model %in% expr.models) %>% 
  ggplot(aes(x = peptides, y = detected_ligands, color = model))+
  geom_line(size = 1)+
  facet_wrap(~tumor, scales = "free")+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_cumsum_expr_final.pdf"), width = 6, height = 2, scale = 2)
```

```{r plot_cumsum_comb, fig.width=12, fig.asp=.4}
XGB.results %>% 
  filter(model %in% comb.models) %>% 
  ggplot(aes(x = peptides, y = detected_ligands, color = model))+
  geom_line(size = 1)+
  facet_wrap(~tumor, scales = "free")+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_cumsum_comb_final.pdf"), width = 6, height = 2, scale = 2)
```


## Area under the curve

Or calculate the AUC of these curves as a summary metric and plot that. Used the `area_under_curve()` function from the [bayestestR](https://easystats.github.io/bayestestR/index.html) package. 
First check maximum AUC to normalize between 0 and 1.

```{r plot_auc_all, fig.width=10, fig.asp=.45}
XGB.results %>% 
  group_by(tumor, model) %>% 
  summarise(auc = area_under_curve(peptides, peptides)) %>% 
  distinct(auc)

XGB.results %>% 
  group_by(tumor, model) %>% 
  summarise(auc = area_under_curve(peptides, detected_ligands) / 357857.5) %>% 
  ggplot( aes(x = tumor, y = auc, fill = model)) +
  geom_bar(stat = "identity", width=0.75, position=position_dodge(width=0.85), color = "black")+
  ggtitle("AUC in top 0.1%")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_AUC_all_final.pdf"), width = 6, height = 2, scale = 2)
```

Core models:

```{r plot_auc_core, fig.width=10, fig.asp=.45}
XGB.results %>% 
  filter(model %in% core.models) %>% 
  group_by(tumor, model) %>% 
  summarise(auc = area_under_curve(peptides, detected_ligands / 357857.5)) %>% 
  ggplot( aes(x = tumor, y = auc, fill = model)) +
  geom_bar(stat = "identity", width=0.75, position=position_dodge(width=0.85), color = "black")+
  ggtitle("AUC in top 0.1%")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_AUC_core_final.pdf"), width = 6, height = 2, scale = 2)
```

```{r plot_auc_expr, fig.width=10, fig.asp=.45}
XGB.results %>% 
  filter(model %in% expr.models) %>% 
  group_by(tumor, model) %>% 
  summarise(auc = area_under_curve(peptides, detected_ligands / 357857.5)) %>% 
  ggplot( aes(x = tumor, y = auc, fill = model)) +
  geom_bar(stat = "identity", width=0.75, position=position_dodge(width=0.85), color = "black")+
  ggtitle("AUC in top 0.1%")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_AUC_expr_final.pdf"), width = 6, height = 2, scale = 2)
```


```{r plot_auc_comb, fig.width=10, fig.asp=.45}
XGB.results %>% 
  filter(model %in% comb.models) %>% 
  group_by(tumor, model) %>% 
  summarise(auc = area_under_curve(peptides, detected_ligands / 357857.5)) %>% 
  ggplot( aes(x = tumor, y = auc, fill = model)) +
  geom_bar(stat = "identity", width=0.75, position=position_dodge(width=0.85), color = "black")+
  ggtitle("AUC in top 0.1%")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_AUC_comb_final.pdf"), width = 6, height = 2, scale = 2)
```

