---
title: "Check importance XGB models"
author: "Kaspar Bresser"
date: "04/03/2022"

output: 
  github_document:
    toc: true
  html_document: 
    theme: simplex
    highlight: pygments
#    code_folding: show
    self_contained: TRUE
    toc: yes
    toc_float:
      collapsed: no
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      autodep  = TRUE,
                      cache = FALSE,
#                     comment = NA,
                      fig.width = 9,
                      fig.asp = 0.618,
                      fig.align = "center")
```


Applied the analyses below to examine the importance of the various features to the XGBoost model. Will look at both models including the feature library.

Import packages. 


```{r loading}
library(here)
library(caret)
library(xgboost)
library(tidytext)
library(ggpointdensity)
library(gghighlight)
library(ggrepel)
library(tidyverse)
```


Get file names of models using the feature library.

```{r import_models}
files <- list.files(here("XGB_models_final"), pattern = "aff")

files %>% 
  str_extract("\\|.*\\|") %>% 
  str_remove_all("\\|") -> names

files <- set_names(files, names)

files
```

Import the models

```{r import_models2}
files %>% 
  map(~read_rds(here("XGB_models_final", .))) %>% 
  set_names(names) -> XGB.models

XGB.models
```

Plot the model metrics. 

```{r get_metrics}
XGB.models %>% 
  map( "results") %>% 
  reduce(bind_rows) %>% 
  mutate(model = names) -> model.metrics 

model.metrics

model.metrics %>% 
  pivot_longer(8:11, names_to = "metric", values_to = "value") %>%
  mutate(which = case_when(str_detect(metric, "SD") ~ "SD",
                           TRUE ~ "value"),
         metric = str_remove(metric, "SD")) %>% 
  pivot_wider(names_from = which, values_from = "value") -> to.plot

to.plot

model.order <- c("aff_only", "aff_chop_only", "aff_chop_Library", "aff_chop_rna", 
                 "aff_chop_ribo", "aff_chop_rna_ribo", "aff_chop_Library_rna", "aff_chop_Library_ribo",
                 "aff_chop_Library_rna_ribo")

to.plot %>% 
  mutate(model = as_factor(model),
         model = fct_relevel(model, model.order)) -> to.plot
```

```{r plot_metrics, fig.width=8, fig.asp=.35}
ggplot(to.plot, aes(x=model, y=value, fill=model)) + 
  geom_bar(stat="identity", color="black", position=position_dodge(), width = .75) +
  geom_errorbar(aes(ymin=value, ymax=value+SD), width=.2,
                 position=position_dodge(.9))+
  facet_wrap(~metric, scales = "free_y", nrow = 1)+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_model_metrics.pdf"), width = 90, height = 40, units = "mm", scale = 4)
```

Get the importance scores for the library model.

```{r get_importance}
importance.scores <- varImp(XGB.models[["aff_chop_Library"]] )


importance.scores$importance %>% 
  as_tibble(rownames = "feature") %>% 
  dplyr::filter(Overall != 0) %>% 
  mutate(log10.importance = log10(Overall)) -> importance.scores
```

Plot the scores, and highlight the top 20.

```{r importance_plots, fig.width=5, fig.asp=.9}

importance.scores %>% 
  mutate(feature = reorder(feature, log10.importance)) %>% 
ggplot(aes(x = feature, y = log10.importance, label = feature))+
    geom_point()+
    theme_classic()+
    geom_point(data = slice_max(importance.scores, log10.importance, n = 20 ), color = "darkred")+
    geom_text_repel(data = slice_max(importance.scores, log10.importance, n = 20 ), 
                    box.padding = 1.4, max.overlaps = 25, color = "darkred", size = 4)+
    theme(axis.text.x=element_blank(),
          axis.ticks.x=element_blank(), panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_importance_pointlabel.pdf"), width = 75, height = 40, units = "mm", scale = 4)
```



Get the importance scores from the randomforest model, and select the top 20 of each class. 

```{r}
read_rds(here("Output", "RF_per_tumor_new.rds")) %>% 
  mutate(importance = map(rf.models, varImp),
         importance = map(importance, "importance"),
         importance = map(importance, as_tibble, rownames = "feature")) %>% 
  select(tumor, classes, importance) %>% 
  unnest(importance ) %>% 
  select(-no) %>% 
  rename(importance = yes) -> rf.importance.table

rf.importance.table %>% 
  group_by(classes, feature) %>% 
  summarise(score = sum(importance)) %>% 
  inner_join(importance.scores) %>% 
  slice_max(score, n = 10) -> top.important
```

Reformat, and plot, highlighting the features that were in the top 20 of the randomforest models. 

```{r plot_classes, fig.width=10}
importance.scores %>% 
  left_join(top.important) %>% 
  mutate(classes = replace_na(classes, "none"),
         feature.label = str_remove_all(feature, "UTR5|UTR3|CDS|count|hsa|\\."),
         feature.label = str_replace_all(feature.label, "_", " "),
         feature = reorder(feature, log10.importance)) -> to.plot

to.plot

ggplot(to.plot, aes(x = feature, y = log10.importance, label = feature.label))+
  geom_point(size = NA)+
  geom_point(data = to.plot %>% filter(classes == "none"), color = "darkgrey")+
  geom_point(data = to.plot %>% filter(classes != "none"), aes(color = classes))+
  geom_text_repel(data = to.plot %>% filter(classes != "none"), aes(color = classes), 
                    box.padding = 1.4, max.overlaps = 25, size = 4)+
  theme_classic()+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_importance_classes.pdf"), width = 90, height = 55, units = "mm", scale = 4)      
```

```{r fig.width=4}
to.plot %>% 
  mutate(classes = as_factor(classes)) %>% 
  mutate(classes = fct_relevel(classes, "none", "UTR5", "CDS", "UTR3", "hsa", "count")) %>% 
ggplot(aes(x = classes, y = log10.importance, color = classes))+
  geom_boxplot(color = "darkgrey", outlier.shape = NA)+
  geom_jitter(width = .15)+
#  geom_text_repel(data = filter(to.plot, classes != "none") %>% group_by(classes) %>%  slice_max(Overall, n = 3), 
#                  aes(color = classes, label = feature.label), 
#                    box.padding = 1.4, max.overlaps = 25, size = 4)+
  scale_color_manual(values = c("grey", "#a11d21", "#de3c26", "#f7921f", "#f5db11", "#b3d343"))+
  geom_hline(data = filter(to.plot, classes == "none"),
             aes(yintercept = median(log10.importance)), 
             linetype ="dashed" )+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_importance_compare_RF.pdf"), units = "mm", width = 100, height = 45, scale = 1.5)
```





Make a boxplot comparing the importance of the different classes of features.

```{r}
importance.scores %>% 
  filter(!(feature %in% c("affMIN", "chop"))) %>% 
    mutate(classes = case_when(str_detect(feature, "UTR5") ~ "5' UTR",
                                  str_detect(feature, "UTR3") ~ "3' UTR",
                                  str_detect(feature, "count") ~ "PTM",
                                  str_detect(feature, "CDS") ~ "CDS",
                                  str_detect(feature, "hsa") ~ "miR",
                                  TRUE ~ "total")) %>% 
  mutate(feature = reorder(feature, log10.importance)) -> to.plot


ggplot(to.plot, aes(x = classes, y = log10.importance, color = classes))+
  geom_boxplot(color = "darkgrey", outlier.shape = NA)+
  geom_jitter(width = .2)+
  theme_classic()+
  theme(panel.grid.major.y = element_line())

ggsave(here("Figs", "XGB_importance_classes_compared.pdf"), width = 70, height = 55, units = "mm", scale = 4)  

```



New analysis


```{r}
varImp(XGB.models[[2]])[["importance"]] %>% 
  as_tibble(rownames = "feature") %>% 
  mutate(model = "full") -> tmp

varImp(XGB.models[[4]])[["importance"]] %>% 
  as_tibble(rownames = "feature") %>% 
  mutate(model = "part") %>% 
  bind_rows(tmp) %>% 
  pivot_wider(names_from = model, values_from = Overall)-> tmp

tmp
```


```{r , fig.width=5}
tmp %>% 
  ggplot(aes(x = log10(part), y = log10(full)))+
  geom_point()+
  theme_classic()+
  theme(panel.grid.major = element_line())
```



